# Random Matrix Theory (RMT)–based random network


> When a numerical matrix is used as input, it is necessary to filter and screen the resulting correlation matrix. Empirical thresholds are commonly adopted, usually between 0.6 to 0.85; however, these thresholds are largely subjective.

> Compared with empirical thresholding, random matrix theory (RMT) provides a data-driven and objective approach for determining correlation thresholds, effectively separating true biological signals from random noise and improving the robustness and comparability of ecological networks.


## Random Matrix Theory (RMT) 



> Load R Package


```{r warning=F, message=F}
library(tidyverse)   # data manipulation / piping utilities
library(ggNetView)   # graph builders + info-extraction helpers
```

----

## Example data

> Example data

```{r}
# Access built-in example datasets in ggNetView

# Relative abundance table of rarefied ASVs or OTUs
data("otu_rare_relative")
dim(otu_rare_relative)
otu_rare_relative[1:5, 1:5]

# Taxonomic annotation table of ASVs or OTUs
data("tax_tab")
dim(tax_tab)
tax_tab[1:5, 1:5]
```


## Compute RMT threshold 


```{r}
# ---- RMT threshold scan #1: WGCNA correlation backend ----
# `ggNetView_RMT()` scans a series of candidate |r| thresholds, builds the
# thresholded correlation matrix at each, computes its eigenvalue spacing
# distribution (NNSD), and picks the threshold at which the NNSD transitions
# from Gaussian Orthogonal Ensemble (signal + noise) to Poisson (noise-free,
# modular network) — this is the RMT-recommended cutoff.
out <- ggNetView_RMT(
  mat              = otu_rare_relative, # input matrix (variables x samples)
  transfrom.method = "none",            # already relative-abundance, no extra transform
  method           = "WGCNA",           # use WGCNA::corAndPvalue for correlation
  cor.method       = "pearson",         # Pearson correlation
  unfold.method    = "gaussian",        # Gaussian KDE unfolding of eigenvalues
  bandwidth        = "nrd0",            # KDE bandwidth rule (stats::density default)
  nr.fit.points    = 20,                # support points (only used if unfold.method = "spline")
  discard.outliers = TRUE,              # IQR-trim extreme eigenvalues before unfolding
  discard.zeros    = TRUE,              # drop all-zero rows/cols after thresholding
  min.mat.dim      = 40,                # early-stop if effective dim drops below 40
  max.ev.spacing   = 3,                 # truncate NNSD tail at spacing = 3
  save_plots       = FALSE,             # don't write diagnostic PNGs to disk
  out_dir          = "RMT_plots",       # (only used when save_plots = TRUE)
  verbose          = TRUE,              # print scan progress to console
  seed             = 1115               # fix RNG for reproducibility
)

# The chosen threshold to feed downstream into build_graph_from_mat(r.threshold = ...)
out$chosen_threshold
```



```{r}
# ---- RMT threshold scan #2: plain `cor()` backend for comparison ----
# Same parameters as `out` above, but with `method = "cor"` (base stats::cor /
# psych::corr.test) instead of WGCNA. Useful as a sanity check — the chosen
# threshold should be very close across the two backends
out2 <- ggNetView_RMT(
  mat              = otu_rare_relative,
  transfrom.method = "none",
  method           = "cor",             # <-- only difference vs. `out`
  cor.method       = "pearson",
  unfold.method    = "gaussian",
  bandwidth        = "nrd0",
  nr.fit.points    = 20,
  discard.outliers = TRUE,
  discard.zeros    = TRUE,
  min.mat.dim      = 40,
  max.ev.spacing   = 3,
  save_plots       = FALSE,
  out_dir          = "RMT_plots",
  verbose          = TRUE,
  seed             = 1115
)

# Compare with out$chosen_threshold — they should be in the same neighborhood.
out2$chosen_threshold
```



## Build graph from matrix based on RMT-threshold

> Build graph object

```{r message=F}
# ---- Construct the tidygraph network using the RMT-selected r threshold ----
# Replaces the usual subjective r-cutoff (e.g. 0.6–0.85) with the data-driven
# value `out$chosen_threshold` returned above.
graph_obj <- build_graph_from_mat(
  mat              = otu_rare_relative,      # variable x sample matrix
  transfrom.method = "none",                 # match the transform used for RMT scan
  r.threshold      = out$chosen_threshold,   # RMT-derived |r| cutoff (objective)
  p.threshold      = 0.05,                   # significance cutoff on adjusted p-values
  method           = "WGCNA",                # correlation backend (must match RMT scan)
  cor.method       = "pearson",              # correlation type
  proc             = "bonferroni",           # multiple-testing correction for p-values
  module.method    = "Fast_greedy",          # community detection algorithm
  node_annotation  = tax_tab,                # taxonomy joined onto each node (by name)
  top_modules      = 15,                     # keep top-15 modules; smaller ones -> "Others"
  seed             = 1115                    # reproducibility
)


# Print to inspect node/edge counts and the columns now available on nodes
# (modularity, modularity2, modularity3, Modularity, Degree, Strength + taxonomy)
graph_obj
```






