# Network layout

> Load R Package


```{r warning=F, message=F}
library(tidyverse)
library(ggNetView)
```


>  Example data

```{r}
# Access built-in example datasets in ggNetView

# Relative abundance table of rarefied ASVs or OTUs
data("otu_rare_relative")
dim(otu_rare_relative)
otu_rare_relative[1:5, 1:5]

# Taxonomic annotation table of ASVs or OTUs
data("tax_tab")
dim(tax_tab)
tax_tab[1:5, 1:5]
```

> Build graph object

```{r message=F}
graph_obj <- build_graph_from_mat(
  mat = otu_rare_relative,
  transfrom.method = "none", #  based your input data
  r.threshold = 0.7,
  p.threshold = 0.05,
  method = "WGCNA",
  cor.method = "pearson",
  proc = "BH",
  module.method = "Fast_greedy",
  node_annotation = tax_tab,
  top_modules = 15,
  seed = 1115
)

graph_obj
```


## Gephi layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "gephi",
  center = F,
  shrink = 0.8,
  layout.module = "random",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```



## Fruchterman–Reingold force-directed layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "fr2",
  center = F,
  shrink = 0.8,
  layout.module = "random",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```


## Diamond layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "diamond",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```



## Kk layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "kk",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```


## Nicley layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "nicely",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```



## Petal layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "petal",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```


## Circle layout

## Petal layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "circle",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```




## Diamond outline layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "diamond_outline",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```

## Grid layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "grid",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```

## Heart_centered layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "heart_centered",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```


## Lgl layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "lgl",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```

## Randomly layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "randomly",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```


## Rectangle layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "rectangle",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```






## Square layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "square",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```



## Star layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "star",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```


## Star_concentric layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "star_concentric",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```


## Stress layout

```{r fig.align='center', fig.width=10, fig.height=10}
p <- ggNetView(
  graph_obj = graph_obj,
  layout = "stress",
  center = F,
  shrink = 0.8,
  layout.module = "adjacent",
  group.by = "Modularity",
  fill.by = "Modularity"
)

p
```







