# Network layout

> Load R Package


```{r warning=F, message=F}
library(tidyverse)   # data manipulation / piping utilities
library(ggNetView)   # graph builders + layout/plotting frontend
```


>  Example data

```{r}
# ---- Built-in example datasets ----

# Relative abundance table of rarefied ASVs/OTUs (rows = ASVs, cols = samples).
data("otu_rare_relative")
dim(otu_rare_relative)
otu_rare_relative[1:5, 1:5]

# Taxonomic annotation; first column must be the ASV/OTU ID.
data("tax_tab")
dim(tax_tab)
tax_tab[1:5, 1:5]
```

> Build graph object

```{r message=F}
# ---- Build a co-occurrence network used by every layout demo below ----
graph_obj <- build_graph_from_mat(
  mat              = otu_rare_relative,  # variables x samples numeric matrix
  transfrom.method = "none",             # already relative abundance, no extra transform
  r.threshold      = 0.7,                # |r| cutoff for keeping an edge
  p.threshold      = 0.05,               # adjusted p-value cutoff
  method           = "WGCNA",            # correlation backend: WGCNA::corAndPvalue
  cor.method       = "pearson",          # Pearson correlation
  proc             = "BH",               # multiple-testing correction (Benjamini-Hochberg)
  module.method    = "Fast_greedy",      # community detection algorithm
  node_annotation  = tax_tab,            # taxonomy joined onto nodes by name
  top_modules      = 15,                 # keep top-15 modules; rest -> "Others"
  seed             = 1115                # fix RNG for reproducibility
)

graph_obj
```



## Full graph layout

### Full network

#### Gephi layout

> layout.module = "random"

```{r fig.align='center', fig.width=8, fig.height=8}
# Gephi-style ForceAtlas2-like layout: good general-purpose layout for
# scale-free / community-rich biological networks.
p <- ggNetView(
  graph_obj     = graph_obj,    # the tbl_graph to draw
  layout        = "gephi",      # global node-positioning algorithm
  center        = F,            # don't pull all modules toward the plot center
  shrink        = 0.8,          # scale each module inward (smaller = tighter modules)
  layout.module = "random",     # how modules are placed relative to each other: "random" / "adjacent" / "order"
  group.by      = "Modularity", # grouping variable used for per-module layout
  fill.by       = "Modularity"  # node fill colour mapped to module ID
)

p
```


> layout.module = "adjacent"

```{r fig.align='center', fig.width=8, fig.height=8}
# Gephi-style ForceAtlas2-like layout: good general-purpose layout for
# scale-free / community-rich biological networks.
p <- ggNetView(
  graph_obj     = graph_obj,    # the tbl_graph to draw
  layout        = "gephi",      # global node-positioning algorithm
  center        = F,            # don't pull all modules toward the plot center
  shrink        = 0.8,          # scale each module inward (smaller = tighter modules)
  layout.module = "adjacent",     # how modules are placed relative to each other: "random" / "adjacent" / "order"
  group.by      = "Modularity", # grouping variable used for per-module layout
  fill.by       = "Modularity"  # node fill colour mapped to module ID
)

p
```



> layout.module = "order"

```{r fig.align='center', fig.width=8, fig.height=8}
# Gephi-style ForceAtlas2-like layout: good general-purpose layout for
# scale-free / community-rich biological networks.
p <- ggNetView(
  graph_obj     = graph_obj,    # the tbl_graph to draw
  layout        = "gephi",      # global node-positioning algorithm
  center        = F,            # don't pull all modules toward the plot center
  shrink        = 1,          # scale each module inward (smaller = tighter modules)
  layout.module = "order",     # how modules are placed relative to each other: "random" / "adjacent" / "order"
  group.by      = "Modularity", # grouping variable used for per-module layout
  fill.by       = "Modularity"  # node fill colour mapped to module ID
)

p
```

#### Fruchterman–Reingold force-directed layout

```{r fig.align='center', fig.width=8, fig.height=8}
# FR2: a force-directed layout (springs pull connected nodes, repulsion pushes
# all others apart). Classic choice; reveals dense vs. sparse regions clearly.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "fr2",        # Fruchterman-Reingold variant
  center        = F,
  shrink        = 0.8,
  layout.module = "random",     # modules placed independently from each other
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```


#### Diamond layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "diamond": modules arranged on a rhombic / diamond-shaped lattice.
# Good for showing many modules side-by-side without overlap.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "diamond",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",   # neighbouring modules placed close together to minimise gaps
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```



#### KK layout

```{r fig.align='center', fig.width=8, fig.height=8}
# Kamada-Kawai force-directed layout: like FR but optimises a stress function;
# tends to give smoother, more even spacing for medium-sized networks.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "kk",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```


#### Nicley layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "nicely": igraph's auto-pick layout — chooses a reasonable algorithm based on
# graph size and structure. Safe default when you don't want to pick yourself.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "nicely",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```



#### Petal layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "petal": each module is laid out as a circular petal radiating from the
# global centre. Visually emphasises modular structure.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "petal",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```


#### Circle layout


```{r fig.align='center', fig.width=8, fig.height=8}
# "circle": all nodes placed on a single ring. Useful for small networks or for
# emphasising edges over node positions.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "circle",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```




#### Diamond outline layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "diamond_outline": diamond layout but nodes pushed toward the module border —
# hollow-looking modules, easier to read edge connections inside each module.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "diamond_outline",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```

#### Grid layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "grid": nodes snapped to a regular rectangular grid.
# Edges may cross a lot, but useful for systematic comparison plots.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "grid",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```

#### Heart_centered layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "heart_centered": stylised heart-shaped frame with nodes pulled toward the
# centre. Mostly cosmetic — handy for talks / decorative figures.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "heart_centered",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```


#### Lgl layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "lgl" (Large Graph Layout): force-directed layout designed for large graphs;
# spreads nodes radially around dense cores. Scales better than FR/KK.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "lgl",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```

#### Randomly layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "randomly": nodes placed at uniformly random positions inside the canvas.
# Baseline / sanity-check layout — not informative on its own.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "randomly",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```


#### Rectangle layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "rectangle": nodes laid out along a rectangular frame.
# Good when you want to free up the centre for annotations / edge bundles.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "rectangle",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```


#### Square layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "square": like rectangle but constrained to a square aspect ratio.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "square",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```



#### Star layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "star": one central node + others arranged as rays radiating outward.
# Best when there is a clear hub node; less useful for community-rich graphs.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "star",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```


#### Star_concentric layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "star_concentric": star layout with multiple concentric rings — nodes ranked
# by some centrality measure end up closer to the centre.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "star_concentric",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```


#### Stress layout

```{r fig.align='center', fig.width=8, fig.height=8}
# "stress": stress-majorisation layout (from graphlayouts). Generally produces
# the cleanest "FR/KK-like" results and is reproducible across runs — a strong
# default for publication figures.
p <- ggNetView(
  graph_obj     = graph_obj,
  layout        = "stress",
  center        = F,
  shrink        = 0.8,
  layout.module = "adjacent",
  group.by      = "Modularity",
  fill.by       = "Modularity"
)

p
```



#### circular_modules_gephi_layout

```{r fig.align='center', fig.width=8, fig.height=8}
p_modules_gephi <- ggNetView(
  graph_obj = graph_obj,
  layout = "circular_modules_gephi_layout",
  center = F,
  shrink = 0.9,
  pointsize = c(1, 5),
  layout.module = "order",
  group.by = "Modularity",
  fill.by = "Modularity",
  anchor_dist = 60,
  mapping_line = T,
  r = 1
)

p_modules_gephi
```


```{r fig.align='center', fig.width=8, fig.height=8}
p_modules_equal_gephi <- ggNetView(
  graph_obj = graph_obj,
  layout = "circular_modules_equal_gephi_layout",
  center = F,
  shrink = 0.9,
  pointsize = c(1, 5),
  layout.module = "order",
  group.by = "Modularity",
  fill.by = "Modularity",
  anchor_dist = 60,
  mapping_line = T,
  r = 1
)

p_modules_equal_gephi
```


#### consensus_module_gephi

```{r fig.align='center', fig.width=8, fig.height=8}
p_consensus_module <- ggNetView(
  graph_obj = graph_obj,
  layout = "consensus_module_gephi",
  center = F,
  r = 2,
  node_add = 7,
  anchor_dist = 50,
  pointsize = c(1,3),
  layout.module = "order"
)

p_consensus_module


p_consensus_module2 <- ggNetView(
  graph_obj = graph_obj,
  layout = "consensus_module_gephi",
  center = F,
  r = 2,
  node_add = 7,
  anchor_dist = 50,
  pointsize = c(1,3),
  add_outer = T,
  expand_outer = 1.3,
  label = T,
  layout.module = "order"
)

p_consensus_module2
```


#### consensus_module_equal_gephi

```{r fig.align='center', fig.width=8, fig.height=8}
p_consensus_module3 <- ggNetView(
  graph_obj = graph_obj,
  layout = "consensus_module_equal_gephi",
  center = F,
  r = 2,
  node_add = 7,
  anchor_dist = 50,
  pointsize = c(1,3),
  layout.module = "order"
)

p_consensus_module3



p_consensus_module4 <- ggNetView(
  graph_obj = graph_obj,
  layout = "consensus_module_equal_gephi",
  fill.by = "Phylum",
  center = F,
  r = 2,
  node_add = 7,
  anchor_dist = 50,
  pointsize = c(1,3),
  layout.module = "order"
)

p_consensus_module4



p_consensus_module5 <- ggNetView(
  graph_obj = graph_obj,
  layout = "consensus_module_equal_gephi",
  fill.by = "Phylum",
  center = F,
  r = 2,
  node_add = 7,
  anchor_dist = 50,
  pointsize = c(1,3),
  add_outer = T,
  expand_outer = 1.3,
  label = T,
  layout.module = "order"
)

p_consensus_module5
```

## Sub graph layout


### bipartite network

```{r}
# Extract the subgraph corresponding to module "1" and "7".
# Internally, get_subgraph():
#   1) splits all nodes by the `Modularity` attribute,
#   2) builds one subgraph per module and stores them in `sub_graph_all`,
#   3) if `select_module` is provided, also returns `sub_graph_select`
#      (the graph filtered to those modules via tidygraph::filter),
#   4) returns `stat_module`: a table of node counts per module across the full graph.
graph_sub <- get_subgraph(graph_obj = graph_obj,
                          select_module = c("1", "7")
                          )

graph_sub$sub_graph_select
```


#### bipartite_layout

```{r fig.align='center', fig.width=8, fig.height=8}
p1 <- ggNetView(
  graph_obj = graph_sub$sub_graph_select,
  layout = "bipartite_layout",
  layout.module = "order",
  center = F,
  scale = T
)

p1
```


#### bipartite_gephi_layout

```{r fig.align='center', fig.width=8, fig.height=8}
p2 <- ggNetView(
  graph_obj = graph_sub$sub_graph_select,
  layout = "bipartite_gephi_layout",
  layout.module = "order",
  node_add = 12,
  anchor_dist = 8,
  pointsize = c(1, 5),
  center = F,
  scale = F
)

p2
```

### tripartite network

```{r}
graph_sub <- get_subgraph(graph_obj = graph_obj,
                          select_module = c("1", "7", "6")
                          )

graph_sub$sub_graph_select
```


#### tripartite_layout

```{r fig.align='center', fig.width=8, fig.height=8}
p3 <- ggNetView(
  graph_obj = graph_sub$sub_graph_select,
  layout = "tripartite_gephi_layout",
  layout.module = "order",
  node_add = 20,
  r = 10,
  pointsize = c(1, 5),
  mapping_line = F,
  anchor_dist = 75,
  center = F,
  scale = F
)

p3
```


#### tripartite_gephi_layout

```{r fig.align='center', fig.width=8, fig.height=8}
p4 <- ggNetView(
  graph_obj = graph_sub$sub_graph_select,
  layout = "tripartite_equal_gephi_layout",
  layout.module = "order",
  node_add = 20,
  r = 10,
  pointsize = c(1, 5),
  mapping_line = F,
  anchor_dist = 100,
  center = F,
  scale = F
)

p4
```

### quadripartite network

```{r}
graph_sub <- get_subgraph(graph_obj = graph_obj,
                          select_module = c("1", "7", "6", "9")
                          )

graph_sub$sub_graph_select
```

#### quadripartite_layout

```{r fig.align='center', fig.width=8, fig.height=8}
p5 <- ggNetView(
  graph_obj = graph_sub$sub_graph_select,
  layout = "quadripartite_gephi_layout",
  layout.module = "order",
  node_add = 15,
  anchor_dist = 8,
  r = 1,
  pointsize = c(1, 5),
  mapping_line = F,
  center = F,
  scale = F
)

p5
```


#### quadripartite_gephi_layout

```{r fig.align='center', fig.width=8, fig.height=8}
p6 <- ggNetView(
  graph_obj = graph_sub$sub_graph_select,
  layout = "quadripartite_equal_gephi_layout",
  layout.module = "order",
  node_add = 15,
  anchor_dist = 8,
  r = 1,
  pointsize = c(1, 5),
  mapping_line = F,
  center = F,
  scale = F
)

p6
```



### pentapartite network

```{r}
graph_sub <- get_subgraph(graph_obj = graph_obj,
                          select_module = c("1", "7", "6", "9", "4")
                          )

graph_sub$sub_graph_select
```


#### pentapartite_layout

```{r fig.align='center', fig.width=8, fig.height=8}
p7 <- ggNetView(
  graph_obj = graph_sub$sub_graph_select,
  layout = "pentapartite_gephi_layout",
  layout.module = "order",
  node_add = 14,
  anchor_dist = 12,
  r = 1,
  pointsize = c(1, 5),
  mapping_line = F,
  center = F,
  scale = F
)

p7
```


#### pentapartite_gephi_layout

```{r fig.align='center', fig.width=8, fig.height=8}
p8 <- ggNetView(
  graph_obj = graph_sub$sub_graph_select,
  layout = "pentapartite_equal_gephi_layout",
  layout.module = "order",
  node_add = 14,
  anchor_dist = 15,
  r = 1,
  pointsize = c(1, 5),
  mapping_line = F,
  center = F,
  scale = F
)

p8
```


### circular network

```{r}
graph_sub <- get_subgraph(graph_obj = graph_obj,
                          select_module = c("1", "7", "6", "9", "4", "2")
                          )

graph_sub$sub_graph_select
```


#### circular_modules_gephi_layout (six group)

```{r fig.align='center', fig.width=8, fig.height=8}
p9 <- ggNetView(
  graph_obj = graph_sub$sub_graph_select,
  layout = "circular_modules_gephi_layout",
  layout.module = "order",
  node_add = 14,
  anchor_dist = 15,
  r = 1,
  pointsize = c(1, 5),
  mapping_line = F,
  center = F,
  scale = F
)

p9
```
#### circular_modules_equal_gephi_layout  (six group)

```{r fig.align='center', fig.width=8, fig.height=8}
p10 <- ggNetView(
  graph_obj = graph_sub$sub_graph_select,
  layout = "circular_modules_equal_gephi_layout",
  layout.module = "order",
  node_add = 14,
  anchor_dist = 15,
  r = 1,
  pointsize = c(1, 5),
  mapping_line = F,
  center = F,
  scale = F
)

p10
```


```{r}
graph_sub <- get_subgraph(graph_obj = graph_obj,
                          select_module = c("1", "7", "6", "9", "4", "2", "11")
                          )

graph_sub$sub_graph_select
```

#### circular_modules_gephi_layout  (seven group)

```{r fig.align='center', fig.width=8, fig.height=8}
p10 <- ggNetView(
  graph_obj = graph_sub$sub_graph_select,
  layout = "circular_modules_gephi_layout",
  layout.module = "order",
  node_add = 14,
  anchor_dist = 15,
  r = 1,
  pointsize = c(1, 5),
  mapping_line = F,
  center = F,
  scale = F
)

p10
```

#### circular_modules_equal_gephi_layout (seven group)

```{r fig.align='center', fig.width=8, fig.height=8}
p11 <- ggNetView(
  graph_obj = graph_sub$sub_graph_select,
  layout = "circular_modules_equal_gephi_layout",
  layout.module = "order",
  node_add = 14,
  anchor_dist = 15,
  r = 1,
  pointsize = c(1, 5),
  mapping_line = F,
  center = F,
  scale = F
)

p11
```
