# Network topology information

> Load R Package


```{r warning=F, message=F}
library(tidyverse)   # data manipulation / piping utilities
library(ggNetView)   # graph builders + info-extraction helpers
```


>  Example data

```{r}
# Access built-in example datasets in ggNetView

# ---- Dataset 1: Relative abundance table of rarefied ASVs/OTUs ----
# Rows = ASVs/OTUs (features / variables), Columns = samples
data("otu_rare_relative")
dim(otu_rare_relative)  # Check matrix dimensions: n_features x n_samples 
otu_rare_relative[1:5, 1:5] # Preview first 5 rows x 5 cols to confirm data layout

# ---- Dataset 2: Taxonomic annotation table for ASVs/OTUs ----
# Rows = ASVs/OTUs (row names should match the abundance table above)
# Columns = taxonomic ranks (Kingdom ~ Species, etc.)
data("tax_tab")
dim(tax_tab)
tax_tab[1:5, 1:5]
```

> Build graph object

```{r}
# ---- Build a co-occurrence network from an abundance matrix ----
# Internal pipeline: transform -> correlation -> p-value adjustment ->
# edge filtering -> community detection -> attach taxonomy.
graph_obj <- build_graph_from_mat(
  mat              = otu_rare_relative,  # variables x samples numeric matrix
  transfrom.method = "none",             # input already relative abundance, no extra transform
  method           = "WGCNA",            # correlation backend: WGCNA::corAndPvalue
  proc             = "bonferroni",       # multiple-testing correction
  r.threshold      = 0.7,                # |r| cutoff for keeping an edge
  p.threshold      = 0.05,               # adjusted p-value cutoff
  node_annotation  = tax_tab,            # taxonomy joined onto each node by name
  top_modules      = 15,                 # keep top-15 modules; rest -> "Others"
  seed             = 1115                # fix RNG for reproducibility
)

graph_obj   # tbl_graph with Modularity / Degree / Strength + taxonomy columns
```


## Get network topology information

> when `mat = NULL`

```{r}
# ---- Compute network topology + robustness for a single graph ----
# When `mat = NULL`, the function still computes all *purely topological*
# metrics from the igraph object, but the following fields are returned as NA
# because they require the original abundance matrix:
#
#   * Cohension_Positive  - needs per-ASV mean abundance to weight positive r
#   * Cohension_Negative  - needs per-ASV mean abundance to weight negative r
#   * Robustness_weight   - abundance-weighted node-removal bootstrap (needs sp.ra)
#   * Robustness_unweight - tied to the same `mat` branch in the code, so also NA
#   * Stability           - mean of the two Robustness_* fields, so also NA
#   * out$Robustness      - the whole 5%-100% removal curve is NA (mean/sd/se)
#
# All other fields (Node, Edge, Degree, Distance, Diameter, Density,
# Transitivity_global/local, Betweenness, Betweenness_edge, Closeness,
# Eigen_centrality, Modularity, K_core_mean/max/min, Network_efficiency,
# Network_info.centrality, Vulenrability) are computed normally from `graph_obj`.
topology_out <- get_network_topology(
  graph_obj        = graph_obj,   # input tbl_graph (required)
  mat              = NULL,        # raw abundance matrix; NULL -> skip cohesion + robustness
  transfrom.method = "none",      # pre-correlation transform (ignored when mat = NULL)
  method           = "WGCNA",     # correlation backend used to rebuild network.raw from `mat`
  proc             = "bonferroni",# multiple-testing correction for p-values (used with `mat`)
  r.threshold      = 0.7,         # |r| cutoff applied when rebuilding network.raw from `mat`
  p.threshold      = 0.05,        # adjusted p-value cutoff applied when rebuilding network.raw
  bootstrap        = 100          # # of random ER networks for `Random_nerwork` column;
                                  # also # of bootstrap iterations for robustness (only when mat != NULL)
)

class(topology_out)

names(topology_out)

topology_out
```

## Get network topology information with matrix

> When a matrix is passed to `mat`, all topology properties are computed.

```{r}
# ---- Compute the FULL set of network topology + robustness metrics ----
# When `mat` is provided, the function recomputes the thresholded correlation
# matrix from `mat` (used as `network.raw`) and ALL topology fields are
# returned with real values - including the ones that are NA when mat = NULL:
#   Cohension_Positive / Cohension_Negative
#   Robustness_weight  / Robustness_unweight
#   Stability
#   out$Robustness     (the 5%-100% node-removal curve)
topology_with_mat <- get_network_topology(
  graph_obj        = graph_obj,         # input tbl_graph (required)
  mat              = otu_rare_relative, # raw abundance matrix used to derive sp.ra and rebuild network.raw
  transfrom.method = "none",            # pre-correlation transform on `mat`; "none" = keep as-is
  method           = "WGCNA",           # correlation backend: WGCNA::corAndPvalue
  proc             = "bonferroni",      # multiple-testing correction
  r.threshold      = 0.7,               # |r| cutoff for keeping an edge in network.raw
  p.threshold      = 0.05,              # adjusted p-value cutoff for keeping an edge
  bootstrap        = 100                # # of bootstrap iterations for robustness AND # of random ER networks for the Random_nerwork column
)


topology_with_mat
```


## Get network topology information from sample 

```{r}
# ---- Per-sample subgraph topology -------------------------------------------
# Step 1: split `graph_obj` into per-sample subgraphs by the abundance matrix
#         (an OTU is "present" in a sample when its abundance > min_abundance).
# Step 2: run the full `get_network_topology()` pipeline on EACH sample subgraph.
# When `mat` is supplied, every topology field (including cohesion, robustness,
# stability) is computed for each sample.
t3_group <- system.time(
  topology_from_sample <- get_sample_subgraph_topology(
    graph_obj        = graph_obj,         # input tbl_graph (the full network)
    mat              = otu_rare_relative, # abundance matrix used for both presence-call AND cohesion/robustness
    transfrom.method = "none",            # pre-correlation transform on `mat`; "none" = keep as-is
    method           = "WGCNA",           # correlation backend: WGCNA::corAndPvalue
    cor.method       = "pearson",         # Pearson correlation
    proc             = "bonferroni",      # multiple-testing correction 
    bootstrap        = 100                # # of bootstrap iterations for robustness; also # of random ER nets for reference
  )
)

topology_from_sample   # one topology table per sample, plus per-sample robustness curves
```





## Get network topology information by parallel

```{r}
# ---- Same as get_network_topology(), but with parallel execution ------------
# Useful when `bootstrap` is large (the random-ER-network reference loop is
# the main bottleneck) or when running on `graph_obj_list` of many networks.
# With `mat = NULL`, cohesion / robustness / stability fields are NA - see
# get_network_topology() docs for the full list of skipped metrics.
t4 <- system.time(
  topology_parallel <- get_network_topology_parallel(
    graph_obj        = graph_obj,
    mat              = NULL,           # no abundance matrix -> cohesion + robustness = NA
    transfrom.method = "none",         # ignored when mat = NULL
    method           = "WGCNA",        # ignored when mat = NULL (no correlation rebuild)
    cor.method       = "pearson",      # ignored when mat = NULL
    proc             = "bonferroni",   # ignored when mat = NULL
    bootstrap        = 100,            # # of random ER networks for the Random_nerwork column
    parallel         = TRUE,           # enable parallel backend
    n_workers        = 2               # # of parallel workers (keep <= cores; 2 is safe under R CMD check)
  )
)

topology_parallel
```

## Get network topology information with matrix by parallel

```{r}
# ---- Parallel + abundance matrix -> FULL topology panel ---------------------
# Same as above, but `mat` is now provided, so every topology field gets a
# real value (cohesion, robustness, stability included).
t5 <- system.time(
  topology_parallel_with_mat <- get_network_topology_parallel(
    graph_obj        = graph_obj,
    mat              = otu_rare_relative, # abundance matrix -> enables cohesion + robustness
    transfrom.method = "none",            # pre-correlation transform on `mat`
    method           = "WGCNA",           # correlation backend used to rebuild network.raw from `mat`
    cor.method       = "pearson",         # Pearson correlation
    proc             = "bonferroni",      # multiple-testing correction
    bootstrap        = 100,               # bootstrap iterations for robustness + # of random ER reference nets
    parallel         = TRUE,              # enable parallel backend
    n_workers        = 2                  # # of parallel workers
  )
)

topology_parallel_with_mat
```








## Integrated Value of Influence (IVI)

```{r}
# ---- Compute IVI (Salavaty et al. 2020) per node ----------------------------
# IVI integrates 6 centrality measures across 3 scales (local / semi-local /
# global) into a single robust influence score. Thin wrapper over
# influential::ivi() — that package must be installed.
obj_aug <- get_node_ivi(
  graph_obj = graph_obj,   # input tbl_graph
  weights   = NULL,        # NULL = unweighted IVI; pass "weight" to use the graph's `weight` edge attribute
  mode      = "all",       # edge mode; "all" is the right choice for undirected graphs
  directed  = FALSE,       # treat the graph as undirected
  d         = 3,           # search radius for the collective-influence component (paper recommends 3)
  scale     = "z-scale",   # "range" [1,100] for single-network view, "z-scale" for cross-network / thresholding (z > 1.645), "none" for raw
  ncores    = 1            # parallel cores for the cluster-rank step; 1 is the safe default
)


obj_aug         # tbl_graph with one extra node column: IVI


class(obj_aug)  # still a tbl_graph - all other vertex/edge columns are preserved


# ---- Top-10 most influential nodes ------------------------------------------
# Activate the node table, convert to tibble, sort by IVI descending, take 10.
obj_aug %>%
  tidygraph::activate(nodes) %>%
  tidygraph::as_tibble() %>%
  dplyr::arrange(desc(IVI)) %>%
  dplyr::slice_head(n = 10)
```
