# Multi-Network and comparison

> Load R Package


```{r warning=F, message=F}
library(tidyverse)   # data manipulation / piping utilities
library(ggNetView)   # graph builders + info-extraction helpers
```


>  Example data

```{r}
# Access built-in example datasets in ggNetView

# ---- Dataset 1: Relative abundance table of rarefied ASVs/OTUs ----
# Rows = ASVs/OTUs (features / variables), Columns = samples
data("otu_rare_relative")
dim(otu_rare_relative)  # Check matrix dimensions: n_features x n_samples 
otu_rare_relative[1:5, 1:5] # Preview first 5 rows x 5 cols to confirm data layout

# ---- Dataset 2: Taxonomic annotation table for ASVs/OTUs ----
# Rows = ASVs/OTUs (row names should match the abundance table above)
# Columns = taxonomic ranks (Kingdom ~ Species, etc.)
data("tax_tab")
dim(tax_tab)
tax_tab[1:5, 1:5]


# ---- Dataset 3: Sample metadata ----
# Rows = samples (sample IDs should match colnames of `otu_rare_relative`)
# Columns = experimental grouping variables (e.g. treatment, time point, site)
data("otu_sample")
otu_sample                    # print the full sample-info table (small enough to view directly)
```

## Comparison of multi-sample networks

### layout = "gephi", scale = FALSE

```{r fig.align='center', fig.width=8, fig.height=8}
# ---- Multi-network "linked" view (no per-group scale normalisation) ----
# `ggNetView_multi_link()` builds ONE network per group (defined by `group_info`),
# lays each one out independently, and places all sub-networks on a circular
# anchor frame so that cross-group edges/relationships can be visually compared.
out1 <- ggNetView_multi_link(
  mat              = otu_rare_relative, # variables x samples abundance matrix
  group_info       = otu_sample,        # sample metadata; defines how samples are split into groups
  transfrom.method = "none",            # pre-correlation transform on `mat`; "none" = keep as-is
  r.threshold      = 0.7,               # |r| cutoff for keeping an edge
  p.threshold      = 0.05,              # adjusted p-value cutoff
  method           = "WGCNA",           # correlation backend: WGCNA::corAndPvalue
  cor.method       = "pearson",         # Pearson correlation
  proc             = "BH",              # multiple-testing correction (Benjamini-Hochberg)
  module.method    = "Fast_greedy",     # community detection per sub-network
  layout.module    = "adjacent",        # neighbouring modules placed close together
  center           = T,                 # pull a node toward the centre of each sub-layout
  top_modules      = 15,                # keep top-15 modules per sub-network; rest -> "Others"
  layout           = "gephi",           # global node-positioning algorithm per sub-network
  shrink           = 0.5,               # shrink each sub-network inward (smaller = more compact)
  scale            = F,                 # do NOT normalise each group to a common scale -> sub-networks keep their native size differences
  jitter           = T,                 # jitter node positions slightly to reduce overlap
  jitter_sd        = 0.3,               # SD of the jitter (bigger -> more spread)
  anchor_dist      = 30,                 # distance between groups on the outer anchor circle
  seed             = 1115,              # fix RNG for reproducibility
  orientation      = "up",              # frame orientation; one of "up"/"down"/"left"/"right"
  angle            = 0                  # additional rotation angle (degrees)
)

p1 <- out1[["p"]]   # the assembled ggplot object
p1

out1$info           # auxiliary table: per-group node/edge counts, module info, etc.
```
### layout = "gephi", scale = TRUE

```{r fig.align='center', fig.width=8, fig.height=8}
####----Plot2----####
# ---- Same setup but with `scale = TRUE` -> each group normalised to a common size ----
# Useful when groups have very different network sizes and you want them
# visually comparable side-by-side. With `scale = TRUE` the native size
# differences are removed, so jitter_sd is also reduced (0.01) accordingly.
out2 <- ggNetView_multi_link(
  mat              = otu_rare_relative,
  group_info       = otu_sample,
  transfrom.method = "none",
  r.threshold      = 0.7,
  p.threshold      = 0.05,
  method           = "WGCNA",
  cor.method       = "pearson",
  proc             = "BH",
  module.method    = "Fast_greedy",
  layout.module    = "adjacent",
  center           = T,
  top_modules      = 15,
  layout           = "gephi",
  shrink           = 0.5,
  scale            = T,                 # normalise each group to a comparable coordinate scale
  jitter           = T,
  jitter_sd        = 0.01,              # smaller jitter — sub-networks are already on a unified scale
  anchor_dist      = 1,
  seed             = 1115,
  orientation      = "up",
  angle            = 0
)


p2 <- out2[["p"]]
p2
```



> add_group_outer = TRUE


```{r fig.align='center', fig.width=8, fig.height=8}
####----Plot2----####
# ---- Same setup but with `scale = TRUE` -> each group normalised to a common size ----
# Useful when groups have very different network sizes and you want them
# visually comparable side-by-side. With `scale = TRUE` the native size
# differences are removed, so jitter_sd is also reduced (0.01) accordingly.
out3 <- ggNetView_multi_link(
  mat              = otu_rare_relative,
  group_info       = otu_sample,
  transfrom.method = "none",
  r.threshold      = 0.7,
  p.threshold      = 0.05,
  method           = "WGCNA",
  cor.method       = "pearson",
  proc             = "BH",
  module.method    = "Fast_greedy",
  layout.module    = "adjacent",
  center           = T,
  top_modules      = 15,
  layout           = "gephi",
  shrink           = 0.5,
  scale            = T,                 # normalise each group to a comparable coordinate scale
  jitter           = T,
  jitter_sd        = 0.01,              # smaller jitter — sub-networks are already on a unified scale
  anchor_dist      = 1,
  seed             = 1115,
  orientation      = "up",
  angle            = 0,
  add_group_outer = T
)


p3 <- out3[["p"]]
p3
```

### layout = "circular_modules_equal_gephi_layout"

```{r fig.align='center', fig.width=8, fig.height=8}
# ---- Circular-modules layout: nodes within each module on equal arcs ----
# Better than "gephi" when you want module structure to dominate the picture.
# `link_level = "Node"`: cross-group links are drawn between individual NODES.
out_multi_link <- ggNetView_multi_link(
  mat                 = otu_rare_relative,
  group_info          = otu_sample,
  transfrom.method    = "none",
  r.threshold         = 0.7,
  p.threshold         = 0.05,
  method              = "WGCNA",
  cor.method          = "pearson",
  proc                = "BH",
  module.method       = "Fast_greedy",
  layout              = "circular_modules_equal_gephi_layout", # per-sub-network layout
  layout.module       = "order",        # modules placed in a fixed order (not random/adjacent)
  top_modules         = 15,
  scale_groups        = T,              # normalise each group to a comparable scale
  add_outer           = "circle",       # outer frame style; "circle" / "manual" / etc.
  r                   = 5,              # radius of each per-group sub-network
  jitter              = F,              # don't jitter nodes
  anchor_dist         = 1,              # distance between groups on the anchor frame
  layout_anchor_dist  = 150,            # anchor frame size in the per-sub-network layout
  seed                = 1115,
  dropOthers          = T,              # drop the "Others" module before plotting
  link_level          = "Node"          # connect cross-group items at NODE level
)

p_multi_link <- out_multi_link[["p"]]
p_multi_link
```



### layout = "circular_modules_equal_gephi_layout"

> link_level = "Node"

```{r fig.align='center', fig.width=8, fig.height=8}
# ---- Duplicate of the previous chunk (kept here for reference) ----
out_multi_link <- ggNetView_multi_link(
  mat                 = otu_rare_relative,
  group_info          = otu_sample,
  transfrom.method    = "none",
  r.threshold         = 0.7,
  p.threshold         = 0.05,
  method              = "WGCNA",
  cor.method          = "pearson",
  proc                = "BH",
  module.method       = "Fast_greedy",
  layout              = "circular_modules_equal_gephi_layout",
  layout.module       = "order",
  top_modules         = 15,
  scale_groups        = T,
  add_outer           = "circle",
  r                   = 5,
  jitter              = F,
  anchor_dist         = 1,
  layout_anchor_dist  = 150,
  seed                = 1115,
  dropOthers          = T,
  link_level          = "Node"
)

p_multi_link <- out_multi_link[["p"]]
p_multi_link
```





### layout = "circular_modules_equal_gephi_layout"

> link_level = "Module&Node2"

```{r fig.align='center', fig.width=8, fig.height=8}
# ---- Cross-group links at "Module & Node2" level ----
# Instead of drawing one line per shared node, draw bundled links between
# modules AND between the corresponding nodes — useful when there are many
# shared items and per-node links become visually cluttered.
out_multi_link2 <- ggNetView_multi_link(
  mat                 = otu_rare_relative,
  group_info          = otu_sample,
  transfrom.method    = "none",
  r.threshold         = 0.7,
  p.threshold         = 0.05,
  method              = "WGCNA",
  cor.method          = "pearson",
  proc                = "BH",
  module.method       = "Fast_greedy",
  layout              = "circular_modules_equal_gephi_layout",
  layout.module       = "order",
  top_modules         = 15,
  scale_groups        = T,
  add_outer           = "circle",
  r                   = 5,
  jitter              = F,
  anchor_dist         = 1,
  layout_anchor_dist  = 150,
  seed                = 1115,
  dropOthers          = T,
  link_level          = "Module&Node2", # bundled module-level + node-level cross-group links
  add_group_outer     = T               # draw a translucent boundary around each group
)
p_multi_link2 <- out_multi_link2[["p"]]
p_multi_link2
```


### add group compare

```{r fig.align='center', fig.width=8, fig.height=8}
# ---- Add cross-group comparison brackets + heavy styling -----------------
# This example shows almost every styling knob the function exposes:
# curved cross-group links, group-comparison annotations, per-group ring
# colours, custom module/node link palettes, etc.
out_multi_link3 <- ggNetView_multi_link(
  mat                          = otu_rare_relative,
  group_info                   = otu_sample,
  transfrom.method             = "none",
  r.threshold                  = 0.7,
  p.threshold                  = 0.05,
  method                       = "WGCNA",
  cor.method                   = "pearson",
  proc                         = "BH",
  module.method                = "Fast_greedy",
  layout                       = "circular_modules_equal_gephi_layout",
  layout.module                = "order",
  top_modules                  = 15,
  scale_groups                 = T,
  add_outer                    = "circle",
  r                            = 5,
  jitter                       = F,
  anchor_dist                  = 1,
  layout_anchor_dist           = 150,
  seed                         = 1115,
  dropOthers                   = T,
  link_level                   = "Module&Node2",
  link_curve                   = T,             # draw cross-group links as curves (not straight)
  link_curve_mode              = "cross",       # curve geometry: "cross" bends links across the centre
  link_curve_adaptive_range    = c(0.7, 1.5),   # min/max curvature; adapts to link length
  link_curve_adaptive_bins     = 7,             # # of bins used to map link length -> curvature
  comparisons                  = T,             # draw group-comparison brackets (like ggpubr style)
  comparisons_groups           = list(          # which group pairs to highlight
    c("WT", "KO"),
    c("WT", "OE")
  ),
  order                        = c("WT", "KO", "OE"),   # MUST contain ALL unique groups in `group_info$Group`
  add_group_outer              = T,             # draw a coloured ring around each group
  add_group_outer_color        = c("steelblue", "coral", "forestgreen"),       # ring border colours (one per group)
  add_group_outer_fill         = c("#d0d1e6", "#fee391", "#e5f5e0"),           # ring fill colours (one per group)
  add_group_outer_fill_alpha   = 0.3,           # ring fill transparency
  add_group_outer_linewidth    = 1,             # ring border line width
  link_color_module            = c("#de2d26", "#756bb1"),   # palette for MODULE-level cross-group links
  link_color_node              = c("#c7e9c0", "#c6dbef"),   # palette for NODE-level cross-group links
  link_linewidth_module        = c(1.5, 1.5),   # line widths for module-level links
  link_linealpha_module        = 0.6,           # alpha for module-level links
  link_linealpha_node          = 1,             # alpha for node-level links
  link_linewidth_node          = 0.5,           # line width for node-level links
  add_group_outer_expand       = 2.5            # how much the outer ring is expanded outward beyond node positions
)

p_multi_link3 <- out_multi_link3[["p"]]
p_multi_link3
```



## multi network (4 groups)

```{r fig.align='center', fig.width=8, fig.height=8}
# ---- Build a 4-group dataset by duplicating the first 6 samples as a fake "ko" group ----
# This is purely for demoing the multi-group layouts below; the new "ko" group
# carries the same abundance values as samples 1:6, just with renamed columns.
otu_rare_relative2 <- otu_rare_relative %>%
  dplyr::bind_cols(otu_rare_relative %>%
                     dplyr::select(1:6) %>%
                     purrr::set_names(paste0("ko", 1:6)))  # 6 new fake samples named ko1..ko6

otu_sample2 <- otu_sample %>%
  dplyr::bind_rows(otu_sample %>%
                     dplyr::slice(1:6) %>%
                     dplyr::mutate(Sample = paste0("ko", 1:6),
                                   Group  = rep("ko", 6))) # 6 new metadata rows, Group = "ko"
```

### group_layout = "row"


```{r fig.align='center', fig.width=8, fig.height=8}
# ---- Place the 4 sub-networks on a single ROW ----
# `group_layout = "row"` puts all groups along one horizontal line (nrow = 1).
out_group4_1 <- ggNetView_multi_link(
  mat                          = otu_rare_relative2,
  group_info                   = otu_sample2,
  transfrom.method             = "none",
  r.threshold                  = 0.7,
  p.threshold                  = 0.05,
  method                       = "WGCNA",
  cor.method                   = "pearson",
  proc                         = "BH",
  module.method                = "Fast_greedy",
  layout.module                = "random",
  center                       = T,
  top_modules                  = 10,
  layout                       = "gephi",
  group_layout                 = "row",         # group-frame layout: "row" / "square" / "triangle_down" / "snake_vertical" / ...
  nrow                         = 1,             # # of rows used when `group_layout = "row"` / grid-like layouts
  shrink                       = 0.9,
  scale_groups                 = T,
  add_outer                    = "manual",      # outer frame drawn manually (uses `r`)
  r                            = 12.5,          # radius of each per-group sub-network
  jitter                       = T,
  jitter_sd                    = 0.01,
  anchor_dist                  = 1.25,
  layout_anchor_dist           = 175,
  seed                         = 1115,
  orientation                  = "up",
  dropOthers                   = T,
  angle                        = 0,
  node_add                     = 10,            # extra padding (in coord units) added around each sub-network
  pointsize                    = c(1, 3),       # node-size range (min, max) mapped from a node-level numeric (e.g. Degree)
  link_level                   = "Module&Node2",
  comparisons                  = T,
  comparisons_groups           = list(          # group pairs to bracket; all names MUST be valid groups
    c("WT", "KO"),
    c("KO", "ko"),
    c("ko", "OE")
  ),
  order                        = c("WT", "KO", "ko", "OE"),  # MUST list ALL 4 unique groups
  add_group_outer              = T,
  add_group_outer_fill_alpha   = 0.3,
  add_group_outer_linewidth    = 1,
  link_linealpha_module        = 0.6,
  link_linewidth_node          = 0.5,
  add_group_outer_expand       = 2.5
)

p_group_4_1 <- out_group4_1[["p"]]
p_group_4_1
```

### group_layout = "square"

```{r fig.align='center', fig.width=8, fig.height=8}
# ---- Place the 4 sub-networks on a 2x2 SQUARE grid ----
# `group_layout = "square"` arranges groups in a square frame (4 corners for 4 groups).
out_group4_2 <- ggNetView_multi_link(
  mat                          = otu_rare_relative2,
  group_info                   = otu_sample2,
  transfrom.method             = "none",
  r.threshold                  = 0.7,
  p.threshold                  = 0.05,
  method                       = "WGCNA",
  cor.method                   = "pearson",
  proc                         = "BH",
  module.method                = "Fast_greedy",
  layout.module                = "order",
  center                       = T,
  top_modules                  = 10,
  layout                       = "circular_modules_equal_gephi_layout",
  group_layout                 = "square",      # 2x2 grid for 4 groups
  scale_groups                 = T,
  add_outer                    = "circle",
  r                            = 12.5,
  jitter                       = T,
  jitter_sd                    = 0.01,
  anchor_dist                  = 1.25,
  layout_anchor_dist           = 175,
  seed                         = 1115,
  orientation                  = "up",
  dropOthers                   = T,
  angle                        = 0,
  node_add                     = 10,
  pointsize                    = c(1, 3),
  link_level                   = "Module&Node2",
  comparisons                  = T,
  comparisons_groups           = list(
    c("WT", "KO"),
    c("WT", "ko"),
    c("WT", "OE")
  ),
  order                        = c("WT", "KO", "ko", "OE"),                # all 4 groups
  add_group_outer              = T,
  add_group_outer_color        = c("steelblue", "coral", "forestgreen", "#756bb1"),  # one border colour per group
  add_group_outer_fill         = c("#deebf7", "#fee391", "#e5f5e0", "#efedf5"),      # one fill colour per group
  add_group_outer_fill_alpha   = 0.4,
  add_group_outer_linewidth    = 1,
  link_color_module            = c("#de2d26", "#756bb1", "#2c7fb8"),       # one colour per comparison pair (module-level)
  link_color_node              = c("#c7e9c0", "#c6dbef", "#fde0dd"),       # one colour per comparison pair (node-level)
  link_linewidth_module        = c(1.5, 1.5, 1.5),
  link_linealpha_module        = 0.6,
  link_linealpha_node          = 1,
  link_linewidth_node          = 0.5,
  add_group_outer_expand       = 2.5
)

p_group_4_2 <- out_group4_2[["p"]]
p_group_4_2
```


## multi network (8 groups)

```{r fig.align='center', fig.width=8, fig.height=8}
# ---- Build an 8-group dataset by repeatedly recycling otu_rare_relative columns ----
# This is purely for demoing the 8-group layouts below; abundance values are
# borrowed from the original matrix and renamed so each fake group has 6 samples.
otu_rare_relative3 <- otu_rare_relative %>%
  dplyr::bind_cols(otu_rare_relative %>% dplyr::select(1:6)   %>% purrr::set_names(paste0("ko",  1:6))) %>%
  dplyr::bind_cols(otu_rare_relative %>% dplyr::select(7:12)  %>% purrr::set_names(paste0("oe",  1:6))) %>%
  dplyr::bind_cols(otu_rare_relative %>% dplyr::select(13:18) %>% purrr::set_names(paste0("wt",  1:6))) %>%
  dplyr::bind_cols(otu_rare_relative %>% dplyr::select(1:6)   %>% purrr::set_names(paste0("aaa", 1:6))) %>%
  dplyr::bind_cols(otu_rare_relative %>% dplyr::select(7:12)  %>% purrr::set_names(paste0("bbb", 1:6)))

otu_sample3 <- otu_sample %>%
  dplyr::bind_rows(otu_sample %>% dplyr::slice(1:6) %>%
                     dplyr::mutate(Sample = paste0("ko",  1:6), Group = rep("ko",  6))) %>%
  dplyr::bind_rows(otu_sample %>% dplyr::slice(1:6) %>%
                     dplyr::mutate(Sample = paste0("oe",  1:6), Group = rep("oe",  6))) %>%
  dplyr::bind_rows(otu_sample %>% dplyr::slice(1:6) %>%
                     dplyr::mutate(Sample = paste0("wt",  1:6), Group = rep("wt",  6))) %>%
  dplyr::bind_rows(otu_sample %>% dplyr::slice(1:6) %>%
                     dplyr::mutate(Sample = paste0("aaa", 1:6), Group = rep("aaa", 6))) %>%
  dplyr::bind_rows(otu_sample %>% dplyr::slice(1:6) %>%
                     dplyr::mutate(Sample = paste0("bbb", 1:6), Group = rep("bbb", 6)))
# Final groups in otu_sample3 (8 total, CASE-SENSITIVE):
#   "WT", "KO", "OE"  (from the original 18 samples)
#   "ko", "oe", "wt", "aaa", "bbb"  (from the bind_rows additions)
```


### group_layout = "triangle_down"


```{r fig.align='center', fig.width=8, fig.height=8}
# ---- Place the 8 sub-networks on a downward-pointing TRIANGLE frame ----
# `group_layout = "triangle_down"` arranges groups along a triangular outline
# with the apex pointing down. Useful for emphasising hierarchy or a "funnel"
# narrative across groups.
out_group4_1 <- ggNetView_multi_link(
  mat                          = otu_rare_relative3,
  group_info                   = otu_sample3,
  transfrom.method             = "none",
  r.threshold                  = 0.7,
  p.threshold                  = 0.05,
  method                       = "WGCNA",
  cor.method                   = "pearson",
  proc                         = "BH",
  module.method                = "Fast_greedy",
  layout.module                = "random",
  center                       = T,
  top_modules                  = 10,
  layout                       = "gephi",
  group_layout                 = "triangle_down",
  nrow                         = 1,
  shrink                       = 0.9,
  scale_groups                 = T,
  add_outer                    = "manual",
  r                            = 5,
  jitter                       = T,
  jitter_sd                    = 0.01,
  anchor_dist                  = 2,
  layout_anchor_dist           = 250,
  seed                         = 1115,
  orientation                  = "up",
  dropOthers                   = T,
  angle                        = 0,
  node_add                     = 10,
  pointsize                    = c(1, 3),
  link_level                   = "Module&Node2",
  comparisons                  = T,
  comparisons_groups           = list(c("WT", "wt"),
                                      c("wt", "KO"),
                                      c("KO", "ko"),
                                      c("ko", "OE"),
                                      c("OE", "oe"),
                                      c("oe", "aaa"),
                                      c("aaa", "bbb")),
  order                        = c("WT", "wt", "KO", "ko", "OE", "oe", "aaa", "bbb"), # all 8 groups in display order
  add_group_outer              = T,
  add_group_outer_fill_alpha   = 0.3,
  add_group_outer_linewidth    = 1,
  link_linealpha_module        = 0.6,
  link_linewidth_node          = 0.5,
  add_group_outer_expand       = 2.5
)

p_group_4_1 <- out_group4_1[["p"]]
p_group_4_1
```



### group_layout = "snake_vertical"


```{r fig.align='center', fig.width=8, fig.height=8}
# ---- Place sub-networks in a vertical SNAKE (zig-zag) frame ----
# `group_layout = "snake_vertical"` runs groups down-up-down... along a vertical
# meander. Useful when you have many groups and want them to fit a tall canvas.
out_group4_1 <- ggNetView_multi_link(
  mat                          = otu_rare_relative3,   # NOTE: this is the 4-group dataset
  group_info                   = otu_sample3,          # NOTE: this is the 4-group metadata
  transfrom.method             = "none",
  r.threshold                  = 0.7,
  p.threshold                  = 0.05,
  method                       = "WGCNA",
  cor.method                   = "pearson",
  proc                         = "BH",
  module.method                = "Fast_greedy",
  layout.module                = "random",
  center                       = T,
  top_modules                  = 10,
  layout                       = "gephi",
  group_layout                 = "snake_vertical",
  nrow                         = 1,
  shrink                       = 0.9,
  scale_groups                 = T,
  add_outer                    = "manual",
  r                            = 10,
  jitter                       = T,
  jitter_sd                    = 0.01,
  anchor_dist                  = 1.25,
  layout_anchor_dist           = 175,
  seed                         = 1115,
  orientation                  = "up",
  dropOthers                   = T,
  angle                        = 0,
  node_add                     = 10,
  pointsize                    = c(1, 3),
  link_level                   = "Module&Node2",
  comparisons                  = T,
  comparisons_groups           = list(c("WT", "wt"),
                                      c("wt", "KO"),
                                      c("KO", "ko"),
                                      c("ko", "OE"),
                                      c("OE", "oe"),
                                      c("oe", "aaa"),
                                      c("aaa", "bbb")),
  order                        = c("WT", "wt", "KO", "ko", "OE", "oe", "aaa", "bbb"), # 8 names, but data only has 4 groups -> will error
  add_group_outer              = T,
  add_group_outer_fill_alpha   = 0.3,
  add_group_outer_linewidth    = 1,
  link_linealpha_module        = 0.6,
  link_linewidth_node          = 0.5,
  add_group_outer_expand       = 2.5
)
p_group_4_1 <- out_group4_1[["p"]]
p_group_4_1
```
